Subject-level developmental relationships

### 2024-08-28 2_developmental_analysis.qmd
switch to percent scores
(rawscore interpretation is form-dependent)

```{r}
library(here)
source(here("helper","common.R"))
d_trial <- readRDS(here("cached_intermediates","1_d_trial"))
d_sub <- readRDS(here("cached_intermediates","1_d_sub"))
```

>>>>

```{r}
suppressPackageStartupMessages(library(here))
source(here("helper","common.R"))
d_trial <- readRDS(here("cached_intermediates","1_d_trial.Rds"))
d_sub <- readRDS(here("cached_intermediates","1_d_sub.Rds"))
```

### 2024-08-28 3_developmental_relationships.qmd
switch to percent scores
(rawscore interpretation is form-dependent)

```{r}
library(here)
source(here("helper","common.R"))
d_sub <- readRDS(here("cached_intermediates","1_d_sub"))
load(here("..", "peekbank-method", "cached_intermediates", "1_cdi_subjects.Rds"))
```

@@

```{r}
d_cdi <- cdi_data |>
  select(administration_id, native_language, measure, language, 
         rawscore) |>
  filter(language == "English (American)", native_language == "eng") |>
  pivot_wider(names_from = "measure", values_from = "rawscore", 
              values_fill = NA, values_fn = mean)
```

@@

```{r}
GGally::ggpairs(select(d_sub_cdi, age, log_age, rt, log_rt, 
                       short_window_accuracy, long_window_accuracy, 
                       short_window_elogit, long_window_elogit, prod, comp), 
        progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

```{r}
GGally::ggpairs(select(d_sub_cdi, log_age, rt,  
                       long_window_accuracy, prod, comp), 
        progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

@@

acc_rt_prc
summary(acc_rt_prc)
ggbiplot::ggbiplot(acc_rt_prc, alpha = .1)

We see that there is a first component with ~60% of variance that is "faster, more accurate, older". Then we see a second component that relates to younger, faster, slightly less accurate. And the final piece is older and less accurate.

These would be interesting components to connect to CDI. In some sense, the claim of some of the Fernald processing corpus is that there is a second principal component here (namely, processing speed) that is meaningful and relates to later learning outcomes.

@@

Probably right now we need to try and connect this to CDIs...

>>>>

```{r}
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(source(here("helper","common.R")))
d_sub <- readRDS(here("cached_intermediates","1_d_sub.Rds"))
load(here("..", "peekbank-method", "cached_intermediates", "1_cdi_subjects.Rds"))
```

@@

```{r}
d_cdi <- cdi_data |>
  select(administration_id, native_language, measure, language, 
         CDI_percent) |> # rawscore is form-dependent
  filter(language == "English (American)", native_language == "eng") |>
  pivot_wider(names_from = "measure", values_from = "CDI_percent", 
              values_fill = NA, values_fn = mean)
```

@@

```{r, fig.width=9, fig.height=9}
GGally::ggpairs(select(d_sub_cdi, age, log_age, rt, log_rt, 
                       short_window_accuracy, long_window_accuracy, 
                       short_window_elogit, long_window_elogit, prod, comp), 
        progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

```{r, fig.width=9, fig.height=9}
GGally::ggpairs(select(d_sub_cdi, log_age, rt,  
                       long_window_accuracy, prod, comp), 
        progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

```{r control-for-age}
d_sub_cdi_mat <- d_sub_cdi |>
  ungroup() |>
  select(age, rt,  long_window_accuracy, prod) |>
  mutate(rt = scale(rt / age)[,1],
         long_window_accuracy = scale(long_window_accuracy / age)[,1], 
         prod = scale(prod / age)[,1]) |>
  select(-age)

acc_rt_prc
summary(acc_rt_prc)
ggbiplot::ggbiplot(acc_rt_prc, alpha = .1)
```

Controlling for age, the first component (~49% of variance) represents high production, fast RT, but low LWL accuracy.
The second component (~31% of variance) captures slow RT, low accuracy, and moderate production, and the third component (~20%) captures fast RT, low accuracy, and low production).

The claim of some of the Fernald processing corpus is that there is a second principal component here (namely, processing speed) that is meaningful and relates to later learning outcomes.

@@

acc_rt_prc
summary(acc_rt_prc)
ggbiplot::ggbiplot(acc_rt_prc, alpha = .1)
```

There is a first component capturing ~50% of the variance that is low accuracy, low production, slower, and younger.
A second component capturing ~30% of the variance is higher accuracy, low production, fast RT, and slightly young.
A third component (~16%) is older, highly accurate, low production, and somewhat slow RT.

@@

The next thing to do is look at any longitudinal data we have, to see how predictive faster LWL at t0 is of higher CDIs at t1.
